### 2021-04-19 draft.Rmd
chunks

---
title: "draft"
author: "Daniel Saggau"
date: "4/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

```{tikz,fig.cap="Structural Causal Model" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {T};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Atomic Intervention", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left =1 cm of A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5 cm)$) (T) {t};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,right=1 cm of A](UC){$U_C$};
\draw[->] (UA) -- (A);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Policy Intervention",  fig.align= "center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!70] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {$T_\pi$};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

# Counterfactuals

Process is described as follows:

(a) Abduction: Cast probability $P(u)$ as conditional probability $P(u|\epsilon)$ 
(b) Action: Exchange $(X = x)$ 
(c) Prediction: Compute $(Y = y)$

# Probabilistic Models vs SCMs

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature CP)                                                                             | $\boldsymbol{\cdot}$ Stable(Nature Eq.)
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific(Invariance of Eq.)
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ no information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors