### 2021-04-22 paper/draft.Rmd
intro

---
title: "A Gentle Introduction into Structural Causal Models"
author: 
- Daniel Saggau $\boldsymbol{\cdot}$ `daniel.saggau@campus.lmu.de`
- Department of Statistics, Ludwig Maximilian University Munich, Germany
date: "April 2021"
output: 
  pdf_document: 
    number_sections: yes
    fig_caption: yes
citation_package: --biblatex
fontsize: 12
linestretch: 1.5
abstract: This paper provides a gentle introduction into causal inference, focusing on structural causal models and respective causal graphs. The first section provides introduces probabilities and will focus on clarifying the difference between observations and causal concepts. The second section discusses the hierachy of causation. The third section dicusses causal graphs.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Topic 1: Structural Causal Models (SCMs)
Structural Causal Models are a general framework to describe data generating mechanisms. 
The values of variables are described as functions of exogeneous and endogeneous variables. 
These mechanisms can also be described with directed graphs. 
Typically one requires for the endogenous variables to fulfill causal sufficiency and the corresponding graph to be acyclic.
Also, one requires mechanisms to be independent.
The student should convey and explain the notation on a number of examples and explain the motivation and role of the aforementioned assumptions.

# Introduction

 |Method          | Action |  Example
|-----------------|---------------------------|------------------------|
| Prediction      | Observation/Co-occurrence | What happened...
| Intervention    | Do-manipulation           | What happens if ...
| Counterfactual  | Hypothetical Realities    | What would have happened if...

Table: Pearl - Hierachy of Causation

## SEM and Structural Causal Model

## Prediction

## Intervention

```{tikz,fig.cap="Structural Causal Model" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {T};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Atomic Intervention", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left =1 cm of A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5 cm)$) (T) {t};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,right=1 cm of A](UC){$U_C$};
\draw[->] (UA) -- (A);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Policy Intervention",  fig.align= "center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!70] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {$T_\pi$};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

## Counterfactuals

Process is described as follows:

(a) Abduction: Cast probability $P(u)$ as conditional probability $P(u|\epsilon)$ 
(b) Action: Exchange $(X = x)$ 
(c) Prediction: Compute $(Y = y)$

## Bayesian Models vs. SCM

## Probabilistic Models vs SCMs

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature CP)                                                                             | $\boldsymbol{\cdot}$ Stable(Nature Eq.)
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific(Invariance of Eq.)
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ no information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors