### 2021-04-22 references/peters_Causality.Rmd
essay

---
title: "peters_elements_causal_inference"
author: "Daniel Saggau"
date: "4/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 1

Probability theory:

Statistical learning: Capacity measures

Peters et al. (2017) suggest that statistical learning theory has a inverse problem: We have data and want to understand the underlying data generating process and respective underlying structures. 
This true underlying data generating process is unknown.
We cannot observe the true DGP based on observational data, alone.
Learning theory proposes the usage of functional specifications of our model to define the true underlying process that is rule based.
These rules are pivotal because they allow us to specify the behavior of the model when exposed to unseen observation data.
To ensure that this specification works, Peters et al. (2017) suggests that statistical learning theory uses the concept of capacity measures. 
If we have high capacity, our functional specification is very applicable to different dataset.
If we have low capacity, our specification is not representative of many datasets.

Apart from having the problem of statistical ill posedness (once because the conditional expectation is undefined for values outside of our sample). 
The second problem is that we usually have no full information for a underlying causal model even when having a fully specified observational distribution.