### 2021-05-03 appendix.Rmd
intro + appendix

---
title: "appendix"
author: "Daniel Saggau"
date: "5/3/2021"
output: pdf_document
---

\footnotesize
 |Method          | Action |  Example | ML Method | 
|---------------------|-------------|-------------------------|-------------------|
| Prediction $P(a|b)$               | Co-occurrence             | What happened...               | Supervised/Unsupervised  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            | Reinforcement Learning      
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if... | Reinforcement Learning / Interpretable ML

Table: Pearls Hierachy of Causation (2009)
\normalsize