### 2021-05-14 presentation_scm.Rmd

---
title: "scm_presentation"
author: "Daniel Saggau"
date: "5/14/2021"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# What is a SCM

- Functional specification of our model including latent factors
- Probabilistic specification + additional knowledge

# SCMs for Fairness in dynamical Systems

- Fair-MDP
- Judea Pearl:do-calculus

# Graphical Illustration

```{tikz,fig.align="center",fig.cap="Probabilistic Model", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\node[my node] (A){A};
\node[my node,right =1 cm of A](C){C};
\node[my node] at ($(A)!0.5!(C)-(0pt,1.5cm)$) (T) {T};
\draw[->] (A) -- (T);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {T};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

# Interventions

- Atomic Intervention 
- Policy Intervention 
- Off-Policy Intervention (model-based; model-free)

# Results Off Policy Intervention - Lambda

* Introduce trade off parameter lambda 
* V of Pi is the overall objective, Pi is our policy and U is utility

$$V_\pi =U - \lambda\delta_{EQOPP}$$

# Causal Hierachy

 |Method          | Action |  Example | Usage | 
|------------------|-------------|--------------------|-------------------|
| Association $P(a|b)$               | Co-occurrence             | What happened...               |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if...           | SCM ,PO

Table: Pearls Hierachy of Causation (2009)

# Exisiting Considerations

- Counterfactuals for Fairness in Dynamical Systems 
- Off-policy estimation (model based (regression) or model free estimation (propensity weight))

# Adding to existing Literature

**Methodological:**

- Cyclic Structural Causal Models with actual reinforcing loops
- Semi-Deterministic SCMs (deterministic -> all variables are known)
-

# Causal Inference and Time 
\tiny

$$
\begin{array}{|l|l|l|l|l|l|}
\hline \text { model } & \begin{array}{l}
\text { predict IID } \\
\text { setting }
\end{array} & \begin{array}{l}
\text { predict under } \\
\text { changing } \\
\text { distributions / } \\
\text { interventions }
\end{array} & \begin{array}{l}
\text { answer } \\
\text { counter- } \\
\text { factual } \\
\text { questions }
\end{array} & \begin{array}{l}
\text { obtain } \\
\text { physical } \\
\text { insight }
\end{array} & \begin{array}{l}
\text { automatically } \\
\text { learn from } \\
\text { data }
\end{array} \\
\hline \begin{array}{l}
\text { mechanistic } \\
\text { model } \\
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & ? \\
\hline \begin{array}{l}
\text { structural } \\
\text { causal model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} & \mathrm{Y} ? ? \\
\hline \begin{array}{l}
\text { causal } \\
\text { graphical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} & \mathrm{N} & \mathrm{Y} ? \\
\hline \begin{array}{l}
\text { statistical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{N} & \mathrm{N} & \mathrm{N} & \mathrm{Y} \\
\hline
\end{array}
$$