### 2021-05-14 references/appendix.Rmd
readme and stuff

---
title: "appendix"
author: "Daniel Saggau"
date: "5/3/2021"
output: pdf_document
---

\footnotesize
 |Method          | Action |  Example | Methods | 
|------------------|-------------|--------------------|-------------------|
| Association $P(a|b)$               | Co-occurrence             | What happened...               |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if...           |SCM

Table: Pearls Hierachy of Causation (2009)
\normalsize