### 2021-05-16 scm.Rmd
scm structure

---
title: "A Gentle Introduction to Structural Causal Models"
author: "Daniel Saggau"
date: "5/11/2021"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation

```{r, out.width= "70%", out.height = "70%", fig.align='center'}
library(knitr)
setwd("~/Desktop/SOSE2021/causal_ml")
include_graphics('./images/correlation.jpg')
```

# Introduction (1)

- Machine learning has provided many insights into different problems 
- One issue is the consideration of 'What are we actually predicting'?
- Mainstream tools are build on association-based learning 
- Associations are not enough for high stake settings 
- In disciplines like psychology or economics people are less interested in associational learning
- We want causation and not correlation

# Introduction (2)

Causal assumptions differentiate causal models from association learning methods.

 | Association-based Concepts |  Causal Concepts
|--------------------------|------------------------|
| Correlation              | Randomization  
| Regression               | Confounding
| Conditional Independence | Disturbance
| Likelihood               | Error Terms      
| Odds Ratio               | Structural Coefficients       
| Propensity Score         | Spurious Correlation

# Table of Contents

* Foundations of SCMs 
  * Assumptions 
  * Comparative Causal Tools 
  * Historical Development 
* Pearl's Causal Hierachy 
  * prediction 
  * intervention 
  * Counterfactuals 
  * Implications 
* Graphical Models 
  * Features 
  * Implications
* Causality and Time

# Assumptions

- Consists of system of equations 
- Assignment equation ':=' rather than regular equation '=' 
- is a nonparametric SEM 
- has functional form rather than using probabilities
- entails features from the PO framework and graphical representation 
- Exogenous factors are part of the model specification

# Assumptions

**Error terms**

Regression: Omittable outside factor 
SCM/SEM; Latent influencial factor that is pivotal for the model specification but not observable

Consists of graph and assignments: 
Baseline:

$$C:= N_c$$
$$E:=f_E(C,N_E)$$

source: @peters_elements_2017

# SCM Applications:

- Flexible simulations for higher order problems (intervention, counterfactual)
- Graphical visualization via directed acyclic graph

# Comparative Causal Tools

# Historical Development

* Path Analysis -> SEM -> SCM

# Fundamental Differences (1)

- conflict whether to use graphs or not 
* A SEM is a parametric specification used in applied sciences (parameters contested)
* A Bayesian causal network is another popular causal model using conditional probabilities and NO functions 
* Differences in performance between BCN and SCM# Performance Evaluation

# Implications

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature CP)                                                                             | $\boldsymbol{\cdot}$ Stable(Nature Eq.)
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific(Invariance of Eq.)
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ no information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors

# Pearls Causal Hierachy

\footnotesize
 |Method          | Action |  Example | Usage | 
|------------------|-------------|--------------------|-------------------|
| Association $P(a|b)$               | Co-occurrence             | What happened...               |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if...           | SCM ,PO

Table: Pearls Hierachy of Causation (2009)

\normalsize

# Graphical Illustration

```{tikz,fig.align="center",fig.cap="Probabilistic Model", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\node[my node] (A){A};
\node[my node,right =1 cm of A](C){C};
\node[my node] at ($(A)!0.5!(C)-(0pt,1.5cm)$) (T) {T};
\draw[->] (A) -- (T);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {T};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

# Causality and Time

- Time in Physical Sciences: Mechanical and exact 
- Time in Social Sciences: Often Vague 
- Regular Time Specification is also more vague 
- To accomodate that issue, research on differential equation based SCMs started