### 2021-05-21 presentation_video.Rmd
SCM

---
title: "Introduction to Structural Causal Models"
author: "Daniel Saggau"
date: "5/11/2021"
citation_package: --biblatex
bibliography: [ref.bib]
#csl: ieee.csl
output: 
  beamer_presentation:
    theme: Madrid
header-includes:
- \definecolor{UBCblue}{rgb}{0.04706, 0.13725, 0.26667}
- \usecolortheme[named=UBCblue]{structure}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation

```{r, out.width= "80%", out.height = "80%", fig.align='center'}
library(knitr)
setwd("~/Desktop/SOSE2021/causal_ml")
include_graphics('./images/correlation.jpg')
```

# Introduction

 | Association-based Concepts |  Causal Concepts
|--------------------------|------------------------|
| Correlation              | Randomization  
| Regression               | Confounding
| Conditional Independence | Disturbance
| Likelihood               | Error Terms      
| Odds Ratio               | Structural Coefficients       
| Propensity Score         | Spurious Correlation

Table: [@pearl2012causal]

# Table of Contents

* Foundations of SCMs 
  * Assumptions 
  * Comparing Causal Tools 
* Pearl's Causal Hierachy 
  * Prediction 
  * Intervention 
  * Counterfactuals 
* Graphical Models 
* Causality and Time

# Foundations of SCMs

- Nonparametric SEM 
- System of equations with functions
- **Assignments** ':=' (non-symmetric) 
- algebraic equation '=' (symmetry)
**Error in Regression:** Omittable outside factor 
**Error in SCM:** Latent (influential) factor that is pivotal for the model

$$C:= N_c$$
$$E:=f_E(C,N_E)$$

source: @peters_elements_2017

# Assumptions in Causality (1)

Independence:

- Noise terms independent $(N_c,N_e)$
- Mechanisms independent (other variables invariant)(local changes)
- Change in distribution stems from change in mechanism 
- Causal Markov Condition

# Assumptions in Causality (2)

*  **SUTVA** 'The treatment that one unit receives does not change the effect of treatment for any other unit.'
* **Consistency** The outcome Y agrees with the potential outcome corresponding to the treatment indicator.' 
* **Ignorability** The potential outcomes are conditionally independent of treatment given some set of de-confounding variables. (perfect RCT)

- First tow hold for Counterfactuals in SCM
- third not testable but can check via backdoor criterion in SCM [@hardtrecht]

# SCM Applications:

- Flexible simulations for higher order problems (intervention, counterfactual)
- Graphical visualization via directed acyclic graph
- Example: SCM for fairness in dynamical system [@creager] : credit loan approval, time allocation, college admission

# Fundamental Differences (1)

- Conflict whether to use **graphs** or not and to use **parameters** in SEM
* SEM is a parametric model used in applied sciences 
* A Bayesian causal network (BCN) use conditional probabilities instead of functions 
* PO is mostly equivalent but has not graphical framework
* SCM and PO are only frameworks that allow for counterfactuals

# Table of Contents

* Foundations of SCMs 
  * Assumptions 
  * Comparing Causal Tools 
* **Pearl's Causal Hierachy**
  * Prediction 
  * Intervention 
  * Counterfactuals 
* Graphical Models 
* Causality and Time

# Pearls Causal Hierachy

\footnotesize
 |Method          | Action |  Example | Usage | 
|------------------|-------------|--------------------|-------------------|
| Association $P(a|b)$               | Co-occurrence             | What happened...               |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if...           | SCM ,PO

Table: Pearls Hierachy of Causation (2009)

\normalsize

# Prediction

- ML, BN and regression are at the lowest level in the causal hierarchy
- Prediction methods demand the least information and depend on association alone
- Association-based methods ignore external changes outside of our data
- Intervention distribution needed for higher level information

# Intervention

- Mathematical Tool: do-calculus
- The do-calculus enables us to study the manipulation of parent nodes
- **Atomic intervention:** where we set a variable to a constant
- **Policy intervention:** we specify a different function for an equation
- CBN , MDP and reinforcement learning model intervention.

# Example Intevention (1)
**Atomic Intervention**:

- replacing function with a constant

$$C_1:= f_{c_1}(p,N_{c_1}) \rightarrow C_1:=600$$
$$C_2:= f_{c_2}(a,N_{c_2})$$
$$E:=f_E(C_1,C_2,N_E)$$

# Example Intevention (2)

**Policy Intervention:**

- replacing function with a different conditional probability

$$C_1:= f_{c_1}(p,N_{c_1}) \rightarrow C_1:=f(\pi)$$
$$C_2:= f_{c_2}(a,N_{c_2})$$
$$E:=f_E(C_1,C_2,N_E)$$

# Counterfactuals

- missing data problem in PO framework

Process is described as follows:

(a) Abduction: Cast probability $P(u)$ as conditional probability $P(u|\epsilon)$ 
(b) Action: Exchange $(X = x)$ 
(c) Prediction: Compute $(Y = y)$

# Table of Contents

* Foundations of SCMs 
  * Assumptions 
  * Comparing Causal Tools 
* Pearl's Causal Hierachy
  * Prediction 
  * Intervention 
  * Counterfactuals 
* **Graphical Models**
* Causality and Time

# Graphical Tools

- Nodes -> Variables (endogenous/exogenous)
- Edges -> relationship (equations)
- Parents/Ancestors/Descendents
- No need to specify exact parametric shape
- Highlight colliders 
- Estimation back door criterion 
- Test theoretical model structure via causal algorithms to detect structure in data (IC/PC Algo.)

# Graphical Illustration - Probabilisitic Model

```{tikz,fig.align="center",fig.cap="Probabilistic Model", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\node[my node] (A){A};
\node[my node,right =1 cm of A](C){C};
\node[my node] at ($(A)!0.5!(C)-(0pt,1.5cm)$) (T) {T};
\draw[->] (A) -- (T);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

# Graphical Illustration - Structural Causal Model

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (T) {T};
\node[my node,right = 1 cm of T, fill=gray!30](C){C};
\node[latent,left = 1cm of T](UT){$U_T$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (T);
\draw[->] (UA) -- (A);
\draw[->] (UT) -- (T);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (T) -- (C);
\end{tikzpicture}
```

# Table of Contents

* Foundations of SCMs 
  * Assumptions 
  * Comparing Causal Tools 
* Pearl's Causal Hierachy
  * Prediction 
  * Intervention 
  * Counterfactuals 
* Graphical Models
* **Causality and Time**

# Causal Modelling with Differential Equations

- Time in Social Sciences: Often Vague 
- Time in Physical Sciences: Mechanical via **Differential equations** 
- dependence on prior time point and change in time contribute to the value at time point

Initial Value: 
$$
\mathbf{x}\left(t_{0}\right)=\mathbf{x}_{0}
$$

Derivative of function x with respect to time t: 
$$
\frac{d \mathbf{x}}{d t}=f(\mathbf{x}), \mathbf{x} \in \mathbb{R}^{d}
$$

Value of Function at time t + dt: 
$$
\mathbf{x}(t+d t)=\mathbf{x}(t)+d t \cdot f(\mathbf{x}(t))
$$

# Graphical Overview

\scriptsize

$$
\begin{array}{|l|l|l|l|l|}
\hline \text { model } & \begin{array}{l}
\text { IID setting }
\end{array} & \begin{array}{l}
\text { changing } \\
\text { distributions } 
\end{array} & \begin{array}{c}
\text { counter- } \\
\text { factual } \\
\text { questions }
\end{array} & \begin{array}{l}
\text { physical } \\
\text { insight }
\end{array} \\
\hline \begin{array}{l}
\text { mechanistic } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} \\
\hline \begin{array}{l}
\text { structural } \\
\text { causal model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} \\
\hline \begin{array}{l}
\text { causal } \\
\text { graphical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} & \mathrm{N} \\
\hline \begin{array}{l}
\text { statistical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{N} & \mathrm{N} & \mathrm{N} \\
\hline
\end{array}
$$
Source: [@peters_elements_2017]
\normalsize

# Extensions

- Mediation Analysis 
- PO-Framework
- Causal Algorithms
- IV-Estimation
- Causal Constraints Model 
- Causal Trees

# Concluding Remarks

- Assosicational learning is easy to model because of lower information neccessary
- but not always appropiate in high stake settings 
- SCM as simulator for causal modelling, entailing a lot of information (DGP, intervention distribution)
- Note: *"Garbage in, Garbage out"*
- computational advantage casting causal model as system of assignments 
- Enables modelling of higher order concepts like counterfactuals 
- Extensions through differential equations for concise modelling of time

\newpage 
# References
\tiny

<div id="refs"></div>
\normalsize
\newpage

# Appendix

\footnotesize

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature CP)                                                                             | $\boldsymbol{\cdot}$ Stable
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific 
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ no information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors

\normalsize