### 2021-05-24 questions.Rmd
Q&A

---
title: "Questions Presentation - Structural Causal Models"
author: "Daniel Saggau"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) What kind of models could the following DAGs be? Briefly justify your choices.

\footnotesize
```{tikz,fig.align="center",fig.cap="Model 1", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\node[my node] (A){A};
\node[my node,right =1 cm of A](C){C};
\node[my node] at ($(A)!0.5!(C)-(0pt,1.5cm)$) (E) {E};
\draw[->] (A) -- (E);
\draw[->] (A) -- (C);
\draw[->] (E) -- (C);
\end{tikzpicture}
```

```{tikz,fig.cap="Model 2", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (E) {E};
\node[my node,right = 1 cm of E, fill=gray!30](C){C};
\node[latent,left = 1cm of E](UE){$U_E$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (E);
\draw[->] (UA) -- (A);
\draw[->] (UE) -- (E);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (E) -- (C);
\end{tikzpicture}
```
\normalsize

2) True or False Questions: Circle the correct answer. Only one answer is correct.

a) The SCM is the same thing as the SEM. There are no differences. 
b) We can derive any potential outcome model based on a structural causal model.
c) The SCM model specification will always outperform other models by design.
d) A Bayesian Causal Network can do anything a Structural Causal Network can do.

3) Write down the respective system of equations for model 2 in question 1.

4) Name two arguments on why we care about SCMs. Also, give two arguments why SCMs might not always be the most suitable model choice.