### 2021-06-11 paper_scm.Rmd
final

---
title: "A Gentle Introduction into Structural Causal Models"
author: 
- Daniel Saggau $\boldsymbol{\cdot}$ `daniel.saggau@campus.lmu.de`
- Department of Statistics, Ludwig Maximilian University Munich, Germany
date: "June 2021"
output: 
  pdf_document:
        number_sections: yes
        fig_caption: yes
link-citations: yes
linkcolor: blue
fontsize: 12
fontfamily: mathpazo 
linestretch: 1.5
citation_package: --biblatex
bibliography: [ref.bib]
#nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Abstract** The interest in understanding relationships of variables beyond co-occurrence has increased the popularity of causal modelling. 
Probabilistic specifications cast a model based on conditional probabilities. 
SCMs cast a model based on asymmetric assignments and extend probabilistic models by specifying the entire data generating process rather than solely utilizing conditional probabilities.
We build a SCM based on a set of assignments and an underlying causal graph. 
A SCM is built on a number of different assumptions namely the Independence of Noise, Autonomy of Mechanisms and Causal Sufficiency.
I also discuss the role of time in causality, focusing on how differential equations for SCMs.
Another difference between SCMs and other model specifications is the ability  of SCMs to address different queries such as *predictions*, *interventions* and *counterfactuals*. 
These queries are part of Pearl's causal hierarchy (2009).
Pearl matches these queries with their respective actions namely *observing*, *doing* and *imagining*. 
I compare the feasibility of addressing these queries.
The insights of this paper can be used as a baseline for subsequent research on structural causal models.

\newpage
\hypersetup{linkcolor=black}
\tableofcontents
\newpage

\hypersetup{linkcolor=blue}

# Introduction

Algorithmic decision making based on co-occurrence is insufficient in high stake settings [@bareinboim2020pearl].
For many problems we want to understand causal relationships between variables.
There are different approaches on how to model causal relationships.
The most popular causal model is the structural causal model or in short SCM.
The geneticist and statistician Sewall Wright introduced the first ancestor of the SCM,the path analysis in the 1910s-1920s [@pearl2009causality; @tarka2018overview].
Path analysis now falls into the broader class of structural equation models (SEM)^[for the specific case of using one variable per indicator.]
SEMs are popular in fields like economics, psychology and sociology [@pearl2009causality].
The SCM is the non-parametric counterpart to structural equation models.
SCMs specify an underlying data generating process without the  burden of creating a correct parametric form^[ Observational data  is seldom consistent. Scholars often question how informative these estimates are. For further information see @hernan2008does and @pearl2012causal].
The SCM is a more flexible version of the SEM.
Researchers on SCMs tried to distance themselves from common practices in structural equation modeling [@pearl2009causality] by using this different name.
Nowadays, scholars often used these terms interchangeably.
The underlying purpose of the SCM is to simulate causal relationships.
This includes underlying latent variables.
Latent noise variables in the SCM are the core of the simulation.
These latent factors define our variables of interest.
In regression analysis we treat noise variables as ignorable factors. 
We assume noise is un-related to our variable of interest.
These perspectives are polar opposites.
SCMs entail endogenous and exogenous variables.
@peters_elements_2017 define endogenous and exogenous variables as follows: *"Endogeneous variables are those that the modeler tries to understand, while exogenous ones are determined by factors outside the model, and are taken as given."*
The basis of these SCMs is a set of functional assignments.
We can use these functions to derive the conditional probabilities for our model [@hardtrecht].
These assignments describe our variables in our model.
Probabilistic models only specify full joint distribution. 
the SCM actually enables the combination of different sources of knowledge.
Sources of information can include observational data but also theory.
In association-based learning we typically only use observational data.
These association-based methods perform poorly in cases of covariate shifts or domain changes.
To account for these changes,  the SCM includes a specification for different interventions (intervention distribution).
Conditional probabilities alone, cannot represent latent variables.
There is no conditional probability in our observational data for unobserved variables [@pearl2009causality].
Hence, we cannot account for changes in our data.
To accommodate existing literature, this paper provides a gentle introduction to SCMs.
The aim of this paper is to provide an intuitive understanding of fundamental concepts within causality.
I focus on the underlying assumptions in SCMs.
This paper focuses on the independence of noise, independence of mechanisms, causal sufficiency (section 2).
Section 3 discusses causal graphs and how we derive underlying graphical representations based on single observational datasets.
Section 4 examines causality and the role of time.
Section 5 studies the causal hierarchy.
I provide considerations for the usage of SCMs in section 6.

# Assumptions in Structural Causal Models

There are various components in SCMs.
When we notate a SCM, we make different assumptions based on the design of these systems and the definition of a SCM.
This sub-section focuses on the mathematical elements in a SCM, the system of equations.
The second sub-section looks at the difference assumptions related to independence.

## Mathematical Components

A SCM contains a (sub-)set of autonomous equations.
These equations are asymmetric assignments.
Regular equations are bidirectional.
Assignments are not bi-directional.
Equations in causal models did not always have a concise notation.
Treating an equation as an algebraic equation led to confusion because those have no causal information.
An algebraic equation would imply that $E=F$ and $C=E$ because the order has no concrete meaning in algebraic equations.
Cause and effect would be interchangeable which is undesirable for causal modelling. 
The initial '=' sign was replaced with the ':=' which is asymmetric and entails causal information.
This misconception has caused a lot of challenges.^[For more information see @pearl2009causality].

There are various definitions for structural causal models. 
Regardless, there is a consensus that a SCM contains two components namely an underlying causal graph and a set of assigments. 
[@peters_elements_2017] define a SCM as follows:

**Definition 1:** Structural Causal Model:

*An SCM* $\mathbb{C}$ *with graph* $C \rightarrow E$ *consists of two assignments*

$$
\begin{array}{l}
C:=N_{C} \\
E:=f_{E}\left(C, N_{E}\right)
\end{array}
$$
*where* $N_{E} \perp \!\!\! \perp N_{C}$ *that is* $N_{E}$ *is independent of* $N_{C}$

In their definition, C is the cause and E is the effect.
$N_C$ is the random noise variable for the cause. 
$N_E$ is the noise variable for the effect variable.

For the subsequent sections, I will use a real world expample.
This example stems from epidemiology.
This study looks at the impact of problem behavior such as smoking on lung cancer. 
Problem behavior is a conceptual term and latent because we cannot observe pure 'problem behaviour' in observational data. 
We can use variables that reflect problem behavior to study problem behaviour. 
One example of problem behavior is smoking.
We can for instance ask participants how frequently they smoke.
As we can see, these functions are driven by underlying latent variables.
These latent factors are the foundation of the structural causal model [@hardtrecht; @pearl2009causality; @pearl2012causal].

\begin{equation}
S:= f_S(U_S)
\end{equation}
\begin{equation}
C:= f_C(S,U_C)
\end{equation}

where: 
$\{S\}$ - Frequency of smoking
$\{C\}$ - Lung cancer

Every structural causal model contains an underlying graphical model [@hardtrecht]. 
This is one important feature that differentiates SCMs from other frameworks^[e.g. the Potential Outcome framework [@pearl2009causality]].

[@pearl2009causality] describes the SCM a process-based tool, because it enables researchers to reflect on their underlying assumptions. 
The SCM requires more assumptions and thought.
Even for a very minimalisitic SCM, we need to define an admissible set of variables, ensure the random noise terms are independent and corroborate that the underlying mechanisms are autonomous. 
By being forced to think about all these steps, SCMs help to avoid poorly specified probabilistic specifications. 
Various research has pointed out examples where modeling without DAGs lead to severe mistakes: 
@hirano2001estimation suggest a method for covariate selection that according to @pearl2009myth favours bias-enhancing features in the propensity score.
Further @bollen2013eight (2013) state that @rosenbaum2002overt and @rubin2007design falsely declared that *'there is no reason to avoid adjustment for a variable describing subjects before treatment'* which also is a severe error.

## Independence of Noise and Mechanism

For the definition of independence, we can use the definition provided by [@peters_elements_2017]:

**Definition 2:** Independence

*"The causal generative process of a system’s variables is composed of autonomous modules that do not inform or influence each other.*
*In the probabilistic case, this means that the conditional distribution of each variable given its causes (i.e., its mechanism) does not inform or influence the other conditional distributions*,
*In case we have only two variables, this reduces to an independence between the cause distribution and the mechanism producing the effect distribution"*.

Suppose we have only two variables smoking and cancer. 
Typically, we can express the full joint probability as either:

\begin{equation}
p(s,c)=p(s|c)p(c)
\end{equation}
\begin{equation}
p(s,c)=p(c|s)p(s)
\end{equation}

In the first example, we define the probability of smoking given one has lung cancer and in the second we look at the conditional probability of lung cancer given smoking.
The central idea is that if we specify the causal effect correctly, we can keep the conditional probability of lung cancer given smoking constant while changing the smoking distribution.
Clearly, these interventions need to be sensible and do not work for both cases.
Writing the probability as smoking behavior given lung cancer and changing lung cancer is an unreasonable intervention.
If we specify the causal structure correctly we can undertake local changes on smoking without changing the conditional probability of lung cancer conditional on smoking.
One example is looking at smoking behavior in a different country (domain shift).
The independence of our mechanisms means that the mechanisms underlying cancer and smoking are independent.
If these terms are independent, we can look at local intervention without rephrasing the entire model and keep the other probabilities invariant.

**Independence of Noise:**

Note, that the way we view noise in SCMs differs from the classical view in regression analysis.
Therefore, I will briefly clarify the differences. 
Noise variables are our latent variables.
These unobserved variables are the parent nodes of our child nodes of interest [@hardtrecht].
In classical regression analysis, such as an ordinary least squares regression model, we employ the Gauss Markov Assumptions to ensure that the estimator is the best linear unbiased estimator (BLUE). 
The exogeneity assumption, one of the five Gauss Markov Assumptions, suggests that the error terms are uncorrelated with our features [@wooldridge2010econometric].
For the SCM, the error terms are driving our features and a pivotal component of the model specification [@hardtrecht].
To ensure that we are correctly specifying mechanisms and our underlying model structure, one needs to ensure independence of these noise terms.
As mentioned in the definition for SCMs, we would notate this independence as follows: $U_{C} \perp \!\!\! \perp U_{E}$ 
where the noise terms of the cause and effect variables are independent.
This is pivotal, because we want to derive conditional probabilities for local changes without having to rephrase other variables. 
There are two related dimensions to consider, namely the informational aspect of independence  and implications for  modularity. 
If the independence conditions holds, and the cause and effect variables are independent, the cause and effect variable do not contain information about each other. 
Modularity [@pearl2009causality] describes the advantage of being able to treat these variables as distinct modules.
This means, even if we have a change in one variable (cause or effect), we can disentangle our variable as unique modular components.
Especially looking at machine learning, domain shift and covariate shift are problematic for classical tools. [@peters_elements_2017]
The central issue of independence of noise variables is that they are latent and henceforth this assumption is untestable.
We can only examine this condition in totality but not in isolation^[via the d-seperation criterion [@hardtrecht]].
Un-testable assumptions are the reason why many statisticians have distanced themselves from causality in the past [@pearl2012causal].
To accommodate the fact that we can never truly ensure that latent variables are independent, @spirtes2000causation proposed the causal sufficiency condition.

**Definition 3:** Causal Sufficiency

*"A set of variables X is usually said to be causally sufficient if there is no hidden common cause $C \notin X$ that is causing more than one variable in X"* [@peters_elements_2017]

By ensuring that this condition holds, we can easier disentangle underlying common causes, avoiding a violation of the independence of our noise terms.
Independence of noise is one form of causal sufficiency [@peters_elements_2017].

# Directed Acyclic Graphs

The most popular graphical model is the directed acyclic graph or in short DAG.
Note that a DAG is a specific graphic model but not every SCM has an underlying DAG.
Every SCM has an underlying graphical model.
A causal graph is a more general graphical model, also including e.g. causal cyclic models.
DAGs are predominately used because they are straightforward but simultaneously only apply if our model is truly directed and acyclic.
A DAG entails nodes (endogenous and exogenous variables) and edges. 
Nodes represent our different variables.
Edges depict the assignment equations.
All edges are directed in the DAG.
If we have some edges without arrows, we call that causal graph semi-directed.
If we have all edges without arrows, we call that causal graph un-directed.
An acyclic graph has no roots that cause itself (directly and indirectly) [@morgan_winship_2014].
This acyclic structure is important for the conditional probabilities [@forre2020causal].
Obtaining conditional probabilities in feedback-loops is challenging [@forre2020causal].
The problem of feedback loops is that we cannot always find unique solutions for the equilibrium state [@peters_elements_2017].
Most DAGs assume the effect is invariant to change over time.[@morgan_winship_2014]
For further information on the role of time in causal modelling, see section 3.

Figure 1 provides a basic structural causal model.
The graph provides an extension of our real world example with another feature namely genetic features.
As one can see, genetic features and smoking are autonous equations and we can derive their conditional distributions in a modular manner. 
Further, the $U_s$, $U_g$ and $U_c$ are the expressive noise variables. 
The square nodes represent the latent variables.
The circle nodes represent the observed variables.
All edges are directed and there are no variables causing itself. 
We are dealing with a DAG.

```{tikz,fig.cap="Structural Causal Model" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (S){S};
\node[latent,left of=S](US){$U_S$};
\node[my node, fill=gray!30] at ($(S)!0.5!(S)-(0pt,1.5cm)$) (G) {G};
\node[my node,right = 1 cm of G, fill=gray!30](C){C};
\node[latent,left = 1cm of G](UG){$U_G$};
\node[latent,right = 1cm of S](UC){$U_C$};
\draw[->] (US) -- (S);
\draw[->] (UG) -- (G);
\draw[->] (UC) -- (C);
\draw[->] (S) -- (C);
\draw[->] (G) -- (C);
\end{tikzpicture}
```

We can obtain a graphical model based on (a single) observational dataset through conditional independence testing.
The graphical approach to recover a graph from an observational dataset requires two assumptions namely jointly independent noise terms (which is assessed via the markov condition) and faithfulness.^[For further information, see [@morgan_winship_2014; @peters_elements_2017]].

Intuitively speaking, this method examines how noise spreads [@peters_elements_2017].
One of the more prominent algorithm to estimate the underlying DAG structure is the pc-algorithm [@kalisch2012causal]. 
Two other methods are ICM^[The intuition here is that the noise terms cultivate the footprint of the assignments. For more information see: @shajarisales2015telling] and using an additive noise model^[This method uses regression and conditional independence testing to disentangle graphical structures. For further information see @mooij2016distinguishing]. 
Graphical models developed a mathematical language to assess certain conditions. 
Two prominent examples are the causal markov condition and the backdoor criterion.
The backdoor criterion allow us to select an admissible subset of variables [@pearl2009causality].
The causal markov condition allows us to test independence in a graph, ensuring us that we are able to utilize the causal factorization [@peters_elements_2017].

# SCMs and Time

In causal modelling we largely ignore time. 
Most SCMs focus on specifying an equilibrium states post-intervention.
Notions of time differ in different disciplines.
Time in social sciences is often vague and less exact.
Natural sciences such as physics treat time in a concise manner, by explicitly including time in models via differential equations [@peters_elements_2017;[@mooij2013ordinary].
This is beneficial if we study non-equilibrium states because we can actual observe the react at different time points rather than only modelling the post intervention outcome.
Dynamical modelling is becoming more relevant in social sciences.
E.g. [@creager] suggest that static concepts in fair-algorithm decision making policies perform very differently if we model these policies as dynamic systems.

We can recast structural causal models based on differential equations, too. 
The first step of casting a SCM with time as a explicit factor is defining an initial state.

\begin{equation}
\mathbf{x}\left(t_{0}\right)=\mathbf{x}_{0}
\end{equation}

Note, that here the initial state is pivotal because there is a dependence on the prior time points.
Subsequently, we need to define our function x at time point t.
We can take a partial derivative with respect to time t of our function x.

\begin{equation}
\frac{d \mathbf{x}}{d t}=f(\mathbf{x}), \mathbf{x} \in \mathbb{R}^{d}
\end{equation}

Thereafter, we combine these elements and look at time t + change in dt:

\begin{equation}
\mathbf{x}(t+d t)=\mathbf{x}(t)+d t \cdot f(\mathbf{x}(t))
\end{equation}

Henceforth, there is a dependence on the initial state and then subsequently we combine this effect with the change.
In our smoking example, one could perhaps think of a situation where we want to examine whether smoking at certain time points is more harmful (impact of smoking during cancer treatment looking at exact time points).

For further context, [@peters_elements_2017] provide a taxonomy of different methods (see table 1).
This table looks at different modelling approaches and how they perform in different settings.

$$\text{Table 1 Source: Peters et al. (2017) , Modelling Taxonomy}$$
$$
\begin{array}{|l|l|l|l|l|}
\hline \text { model } & \begin{array}{l}
\text { IID setting }
\end{array} & \begin{array}{l}
\text { changing } \\
\text { distributions } 
\end{array} & \begin{array}{c}
\text { counter- } \\
\text { factual } \\
\text { questions }
\end{array} & \begin{array}{l}
\text { physical } \\
\text { insight }
\end{array} \\
\hline \begin{array}{l}
\text { mechanistic } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} \\
\hline \begin{array}{l}
\text { structural } \\
\text { causal model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} \\
\hline \begin{array}{l}
\text { causal } \\
\text { graphical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{Y} & \mathrm{N} & \mathrm{N} \\
\hline \begin{array}{l}
\text { statistical } \\
\text { model }
\end{array} & \mathrm{Y} & \mathrm{N} & \mathrm{N} & \mathrm{N} \\
\hline
\end{array}
$$

In table 1 we can see that traditional statistical models fair well in IID settings, where our observations are independent and identically distributed.
Once we encounter a changing distribution such as an covariate shift or a domain shift, these statistical models are not suitable anymore.
Statistical models are unable to answer any counterfactual questions about hypothetical settings outside of our observational data and are unable to provide physical insights because we do not model time as an explicitly. 
Causal graphical models are able to deal with changing distributions and fair well in the IID setting. 
Nevertheless, they are unable to answer counterfactual questions on its own and cannot provide physical insights.
Note that we can derive the underlying causal graphs from a SCM.
This table examines graphical models in isolation apart from the SCM.
SCMs are able to deal with the first three queries, but largely ignore time.
By commonly making assumptions such as that the effect is acyclic, SCMs simplify time and treat focus on equilibrium states.
The mechanistic view, where we model time as an explicit factor, enables us to answer all those queries. 
Note that the growing complexity makes it very challenging to implement but relevant for natural sciences where time is pivotal to disentangle relationships [@mooij2013ordinary; @peters_elements_2017].

# Pearl's Causal Hierachy

@pearl2009causality] introduced the hierarchy of causation to categorize different statistical and causal tools. 
The hierarchy of causation contains (see table 1) three levels, where the high methods on the hierarchy, the more information the method requires.
This section discusses each of these methods and their respective advantages.
I also included the usage of different methods and briefly discuss how they relate to the hierarchy based on [@bareinboim2020pearl].

$$\text{Table 2 Source: Pearl (2009) , Hierarchy of Causation}$$ 
\footnotesize

|Method          | Action |  Example | Usage | 
|------------------|-------------|--------------------|-------------------|
| Association $P(a|b)$               | Co-occurrence             | What happened...               |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation           | What happens if ...            |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals   | What would have happened if...           | SCM ,PO

\normalsize

## Association:

The first level, association, requires the least information.
This query deals with questions like 'what happend?'.
In our smoking example, a possible question is e.g. 'what was the impact of smoking on lung cancer'?
association-based methods are most prevalent and contain the largest class of methods.
Standard statistical tools such as regression analysis, supervised and unsupervised learning and Bayesian Networks all fall into this category [@bareinboim2020pearl].
The underlying action for association is co-occurrence.
As prominently criticized by @bender2021dangers, this reduces the number of questions we can answer, because methods are heavily dependent on the observational data. 
In the context of deep learning and the advancement of natural language processing, @bender2021dangers suggest that many association-based methods results in stochastic parrots as opposed to natural langauge understanding.
Association-based methods ignore external changes outside of our data.
The interventional distribution has information on these external changes.
Note, that the intervention distribution is only defined in high order methods. 
Reiterating our lung cancer example, the full joint distribution looks as follows: $P(c,s)= P(s) \times P(c|s)$

If there are any changes in our distribution, e.g. smoking changes its distribution from $s$ to $s_{new}$ we are unable to accommodate these changes [@peters_elements_2017].
These conditional probabilities are typically not derived from functions $do(S=s)$ in these methods. 
They are based on observational data for the given methods. 
As mentioned in section 1, the amount of things we can derive based on a specific observational dataset is limited [@hernan_causal_nodate; @hernan2008does].
Association-based methods are not equipped to accommodate such modifications.

There are ways to also graphically depict associations-based models. 
One example is the Bayesian Network [@pearl2009causality].
In the probabilistic representation (see figure 2), we ignore latent factors [@creager; @pearl2009causality].
We can see, that this model is technically speaking also a directed acyclic graph.

```{tikz,fig.cap="Probabilistic Model" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (p){S};
\node[my node, fill=gray!30] at ($(p)!0.5!(p)-(0pt,1.5cm)$) (G) {G};
\node[my node,right = 1 cm of G, fill=gray!30](C){C};
\draw[->] (p) -- (C);
\draw[->] (G) -- (C);
\end{tikzpicture}
```

While in this simple 2 feature case, the graphical illustration is rather straight forward. 
But especially in a high dimensional space, these specifications become convoluted.

## Intervention:

Intervention deals with questions like 'what happens if'. 
E.g. in our smoking example, we could look at e.g. 'what happens if people smoke less?' or, 'what happens if people smoke more?'.
This higher order methods opens many different applications and is predominately possible in Causal Bayesian Networks (CBN), Reinforcement Learning (RL) or Markov Decision Processes (MDP).
Note that while CBNs are capable of computing interventions, they are computationally more costly compared to SCMs [@pearl2009causality].
SCMs are built on functions which are inherently better capable of computing changes.
For further informaation on the computational difference, see the appendix. 
For interventions, we can use @pearl2009causality do-calculus.
The do-calculus enables us to study the manipulation of parent nodes.
Instead of merely seeing the co-occurence of variables, we can actively manipulate the conditional distribution of one variable.
Commonly, one could use this method to evaluate different policies [@creager].
In our cancer example one can actively set the smoking behavior to a fixed value or a different conditional probability.
The post intervention joint distribution would look as follows: 
$P_{S=s}(c,g)= P(c|S=s,g) \times P(g)$ 
where $S=s$ is the new probability or atomic value.
There are various types of intervention.
I will illustrate atomic intervention and policy intervention based on our smoking example.

In **atomic intervention,** we set a variable to a constant value.
As one can see in figure 3, c is constant that is not dependent on the latent factor, because in atomic intervention we do not derive the value of c based on the function S.

```{tikz,fig.cap="Atomic Intervention" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!10] (p){c};
\node[my node, fill=gray!30] at ($(p)!0.5!(p)-(0pt,1.5cm)$) (G) {G};
\node[my node,right = 1 cm of G, fill=gray!30](C){C};
\node[latent,left = 1cm of G](UG){$U_G$};
\node[latent,right = 1cm of p](UC){$U_C$};
\draw[->] (UG) -- (G);
\draw[->] (UC) -- (C);
\draw[->] (p) -- (C);
\draw[->] (G) -- (C);
\end{tikzpicture}
```

Further, in mathematical notation the model would change as follows:

\begin{equation}
S:= c
\end{equation}
\begin{equation}
G:= f_G(U_G)
\end{equation}
\begin{equation}
C:= f_C(S,G,U_C)
\end{equation}

In **policy intervention** (see figure 4) we specify a different conditional probability $do(S=s)$ for an equation.
We can derive s from S, because we include information on the intervention distribution in our latent variables.
This information cannot be obtained, if we directly specify our model as conditional probabilities.

```{tikz,fig.cap="Policy Intervention" ,fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!10] (p){s};
\node[latent,left of=p](US){$U_S$};
\node[my node, fill=gray!30] at ($(p)!0.5!(p)-(0pt,1.5cm)$) (G) {G};
\node[my node,right = 1 cm of G, fill=gray!30](C){C};
\node[latent,left = 1cm of G](UG){$U_G$};
\node[latent,right = 1cm of p](UC){$U_C$};
\draw[->] (US) -- (p);
\draw[->] (UG) -- (G);
\draw[->] (UC) -- (C);
\draw[->] (p) -- (C);
\draw[->] (G) -- (C);
\end{tikzpicture}
```

where $s$ is our new conditional probability. 
One of the most prominent ways to estimate casual effects of our treatment is the **average treatment effect**.
The average treatment effect looks at the difference between treatment and control group.
For the case of simplicity in our example, we want to look at whether smoking at all has an impact on lung cancer.
Mathematically, this could as follows:

\begin{equation}
f(C|S=1) - f(C|S=0)
\end{equation}

Where $S = 1$ means the participant smokes and $S=0$ means the participant does not smoke.
Furthermore, it is crucial that we hold all other factors constant (e.g. age, gender, genetic code).

**Confounders:**
Confounding is the circumstance where the observational information (conditional probabilities in our data) and intervention are different.
Essentially, we want to express our do-intervention based on conditional probabilities, but have no direct information in our data.
To still estimate a conditional probability, we can use a mixture of existing conditional probabilities and estimate the desired probability based on the adjustment formula.
For further information, see [@barocas].

## Counterfactuals

Counterfactuals deal with hypothetical situations.
Typically questions are: 'what would have happend if'.
In our smoking case, suppose we want to see what would have happened if we wanted to include a different problem behavior that was not included in the observational data.
Philosophically speaking, we interpret these outcomes as actual outcomes^[In other frameworks such as the Potential outcome framework we interpret them as potential outcomes rather than true outcomes [@hardtrecht]].
The two most prominent models that can obtain counterfactuals are the structural causal model and the potential outcome framework.
This section focuses on counterfactuals in structural causal models. 
For information on how to obtain counterfactuals in the potential outcome framework, see @cunningham2021causal.

SCMs derive Counterfactuals.
We can describe the process as follows:

(a) Abduction: Cast probability $P(u_s)$ as conditional probability $P(u_s|\epsilon)$ 
(b) Action: Exchange $(S = s')$ 
(c) Prediction: Compute $(C = c')$

In the first step, the abduction step, we cast the new conditional probability for our latent factor.
These probabilities is based on the events in our data which are denoted as $\epsilon$.
As one can see, we can only cast the probability for latent factors if we specify them in our model.
Henceforth, models such as the Bayesian Causal Network or a Markov Decision Process cannot provide counterfactuals because they do not include information on these factors.
In the second step, we exchange S for s'.
In the third step, we compute the outcome c' for this hypothetical setting.

Regardless of whether dealing with counterfactuals in SCMs or in the potential-outcome-Framework, we need to ensure three assumptions.
Here, I am using the definitions by [@hardtrecht]:

**Stable Unit Treatment Value Assumption {SUTVA}:**
*"The treatment that one unit receives does not change the effect of treatment for any other unit."*

**Consistency**: 
*"The outcome Y agrees with the potential outcome corresponding to the treatment indicator."*

The first two conditions hold for counterfactuals in structural causal models.
As suggested by [@hardtrecht], the third condition ensures that we are dealing with a perfect randomized controlled trial.

**Ignorability**:
*"The potential outcomes are conditionally independent of treatment given some set of de-confounding variables."*

We can never truly ensure that we are undertaking an experiment under perfect conditions, because we cannot ensure that we include all pivotal variables. 
Therefore, the third condition is an un-testable assumption.
Due to the backdoor criterion, we can nevertheless verify, that our model specification is consistent with this assumption [@hardtrecht].
Note, that this is merely consistency and not a test that allows us to corroborate this assumptions entirely. 
The backdoor criterion is unique to causal graphs.
E.g. the potential outcome framework cannot utilize this test because the framework neglects underlying graphical models.

# Considerations

Structural causal models are data driven models [@hernan_invited_2015]. 
There is a strong dependence on data quality.
In a situation, in which we are certain that the quality of observational data is low, using a SCM is not advisable.
Alternatively, if we have a very strong theoretical understanding, we might be more interested in other modelling approaches such as Markov Decision Processes [@hernan_invited_2015].
Evaluating different sources of information into causal models accordingly is an open problem in causal modelling [@spirtes2010introduction].
If our insights based on data are limited, or perhaps we are not really able to clearly define function relationships beyond conditional probabilities, structural causal models can be problematic.
Furthermore, sometimes association-based knowledge is sufficient.
Not every setting is a high-stake decision making setting.
inevitably, the entry barrier to higher order causal models is higher in terms of required knowledge and resources. 
Not every situation accommodates these conditions nor requires them.
If we have cyclic structures, such as feedback loops, SCM are usable but very complicated.
For further work on this issue see [@mooij2013ordinary; @forre2020causal].

Nevertheless, in a situation where we have high-stake repercussions based on our model, causal modelling has been a complementary tool.
When dealing with policy implementations, that might be controversial or unethical in the real world, simulations are pivotal.
Recently, research on fairness has adopted various concepts and developed new characteristics such as counterfactual fairness [@kusner_counterfactual_2018].
Further fairness research also looks at latent sensitive features such as race and allows us to model bias [@creager].
Another interested area is simulations for situations where our historical data was bias. 
One example is the use-cases by @ensign2018runaway, examining bias in predictive policing based on algorithmic decision making.
Their research suggest that historical bias is reinforced through further racial targeting of drug-related crimes in Oakland based on historical records.
In these cases, we can use SCMs to simulate domain shift and escape reinforcing biases.

# Conclusion

Association-based learning is insufficient to answer advanced queries in high-stake situations. 
Causal Modeling provides the necessary toolkit to deal with questions of  higher order causal understanding. 
The most prominent causal model is the structural causal model, containing mathematical and graph-theoretic elements. 
SCMs are flexible simulators to disentangle causality for higher order queries (e.g. interventions, counterfactuals).
SCMs enable us to consider latent factors, forcing us to re-evaluate existing assumptions in our model.
We make a number of  assumptions when building a structural causal model.
The most important assumption is the independence of mechanisms and independence of noise terms.
Note, that noise terms are the elementary core in SCMs. 
Independence of mechanisms ensures that are mechanisms we can make local changes without re-specifying the entire model.
Independence of Noise ensures that we do not misspecify common causes.
Independence of noise is an un-testable assumption.
We can use the causal sufficiency condition to at least verify our set of variables is consistent with these assumptions.
Every SCM also entails a causal graph.
We can reconstruct graphical models from our observational based on further assumptions.
One can use conditional independence testing and assess assumptions about faithfulness and jointly independent noise terms in the graphical approach. 
These graphs come with a mathematical language to test causal assumptions that are otherwise un-testable and merely assumed to hold true(e.g. backdoor-criterion).
The entry barriers to causal modeling are high .
SCMs demand considerable knowledge about the model.
Further, there are also still many open questions in causal modelling [@spirtes2010introduction].
One example is how to weight different data sources and their relative value.
This issue is of great importance when mixing different data sources.  
Regardless, SCMs provide the building blocks for many different applications.
While their entry requirements are high, the potential of these models is fruitful.
Advances in algorithmic decision making suggest potential, especially looking at the intersection of  causal modeling and machine learning.

\newpage

# References

<div id="refs"></div>
\newpage

# Appendix

$$\text{Table 3: Source: Pearl (2009)}$$

\footnotesize

 |Method         | CBN |  SCM
|----------------|---------------------------|------------------------|
| Prediction     | $\boldsymbol{\cdot}$ Unstable                                                                                        | $\boldsymbol{\cdot}$ Stable                 
|                | $\boldsymbol{\cdot}$ Volatile to parameter changes                                                                   | $\boldsymbol{\cdot}$ More Natural Specification 
|                | $\boldsymbol{\cdot}$ Re-Estimate entire model                                                                        | $\boldsymbol{\cdot}$ Only estimate $\Delta$ CM
|                |                                                                                                                      | 
| Intervention   | $\boldsymbol{\cdot}$ Costly for Non-Markovian Models                                                                 | $\boldsymbol{\cdot}$ Pot. Cyclic Representation
|                | $\boldsymbol{\cdot}$ Unstable(Nature Conditional prob.)                                                                             | $\boldsymbol{\cdot}$ Stable(Nature Equation)
|                | $\boldsymbol{\cdot}$ Only generic estimates($\Delta$ CP)                                                             | $\boldsymbol{\cdot}$ Context specific(Invariance of Equation)
|                |                                                                                                                      | 
|Counterfactuals | $\boldsymbol{\cdot}$ **Impossible**                                                                                  | $\boldsymbol{\cdot}$ Possible
|                | $\boldsymbol{\cdot}$ No information on latent factors($\epsilon$)                                                    | $\boldsymbol{\cdot}$ Inclusion of latent factors

\normalsize